Numerical Stability of Detonations in White Dwarf Simulations

### 2018-07-13 papers/ignition/paper.tex
Strip reaction equations

\section{Nuclear Burning Methodology}
\label{sec:burner}

We first describe the core operation of our burner module, which is implemented in
the \microphysics\ code repository\footnote{\microphysics\ can be obtained at
  \url{https://github.com/starkiller-astro/Microphysics}.}, a collection of microphysical
routines that are compatible with \castro, \maestro, and other \amrex-based astrophysics
codes. We review this because, even though the ideas are not new,
it serves as an update on the state of the \castro\ code since the original release of
the code paper, similar to the goal of \cite{wdmergerI} for the hydrodynamics and gravity
modules.

Given a set of nuclides and the reaction links between them, we consider
how a burning step is performed in our software. The goal is to integrate the
vector ${\bm{Y}} = (Y_1, Y_2, \ldots, Y_n, e, T)$, where $Y_{n} = X_{n} / A_{n}$
is the molar fraction of species $n$, with $X_n$ the mass fraction and $A_n$ the
mass number of that species, $e$ is the energy released during the burn, and
$T$ is the temperature. The equation describing its evolution is given by
\begin{equation}
  \frac{d\bm{Y}}{dt} = f(\mathbf{Y}),
\end{equation}
where the components of the right-hand-side for the species come from the particular
nuclear burning network we are using. The energy $e$ of the zone
will change when the nuclear abundances evolve, according to
\begin{equation}
  \frac{\partial e}{\partial t} = N_A \sum_{n} \frac{\partial Y_{n}}{\partial t} m_{n} c^2,
\end{equation}
where $c$ is the speed of light and $m_n$ is the mass of each nuclide.

For most of the problems of interest to \castro\, we employ a self-heating burn,
where we allow the temperature to evolve in response
to the burning (see\footnote{In the cited paper, a term based on the
thermodynamic chemical potential was included; we now believe
that it is incorrect to include such a term in the burn, since it
automatically sums to zero analytically.} \citet{maestro3}):
\begin{equation}
  \frac{dT}{dt} = \frac{1}{c_V}\frac{\partial e}{\partial t}
\end{equation}
Here $c_V$ is the specific heat at constant volume, which
is provided by the equation of state. During this burn, we re-evaluate $c_V$
by calling the equation of state using the latest value of $(\rho, T)$.
The added expense of the equation of state call (relative to keeping $c_V$ fixed
through the burn) is small in comparison to the more expensive parts of the calculation,
and it can significantly speed up convergence near NSE. Although $T$ evolves during the
burn so that the computation of the reaction rates is accurate, we discard the final value
for $T$ at the end of the burn and recompute a temperature for the zone that is consistent
with its new internal energy according to the equation of state.

In our \microphysics\ repository we provide several software options for
solving a set of coupled stiff ODEs. For this work our primary integrator
is the classic \vode\ integrator \citep{vode}, of which we have a version
compatible with our software interfaces in the \microphysics\ repository.
Another option is an implementation of the well known variable-order Richardson
extrapolation method presented by \citet{stoer:1980}.

>>>>

The numerical simulations in this paper are performed using the reacting hydrodynamics AMR
code \castro\ \citep{castro}. \cite{wdmergerI} described updates to the core hydrodynamic
algorithm since the code's original release. The goal of this paper is to describe several
limitations we found when coupling the hydrodynamics module to the reactions module for detonation
problems, and the changes we made to the \castro\ code to deal with those limitations (though
most of the ideas are applicable to any adaptive mesh-based code). The challenge of accurately
capturing this coupling explains most of the difficulty in obtaining converged detonation
simulations.
